---
title: "Pipes"
author: "Alex Montgomery"
date: "10/03/2022"
output: html_document
---

```{r libraries, include=FALSE}
library(tidyverse) #Load tidyverse
library(ggplot2) #Load ggplot2
library(stargazer) #Load stargazer
```

```{r overwatch logo, out.width = "50%", fig.align = "center", echo=FALSE}
#Load and display pipe fitting jpeg
img1 <- "images/professional-pipe-fitter.jpg"
knitr::include_graphics(img1)
```

## Pipes Dataset
The following dataset that shall be explored contains work orders from a utility company. The data represents jobs performed on pipes belonging to the company.

```{r load data, echo=TRUE, warning=FALSE, include=TRUE}
#Load the data set
dfraw = read.csv("data/raw/example.work.orders.csv")
#Check the data is as expected
head(dfraw, 5)
```

```{r structure data, echo=TRUE, warning=FALSE, include=TRUE}
#Check the structure of the data
str(dfraw)
```

It is important to check the structure of the data to, firstly, get to know the data better, but also to ensure each variable is the correct data type. For example, ensuring a numerical variable hasn't been coded as a string.

## Codebook
Variable Title | Explanation |
|:------:|:------|
JOB_ID | The unique identity of the work order |
JOB_TYPE | The type of job performed |
DEPTH_M | The depth of the pipe (m) |
DIAMETER MM | The diameter of the pipe (mm) |
MATERIAL | The material of the pipe |
LENGTH_M | The length of the pipe (m) |
JOB_COST | The cost of the job (£) |


## Data Cleansing
* Examining the descriptive statistics of a dataset can help identify outliers and data input errors. 
* In this dataset, diameter appears to have an unlikely maximum value of 999999 mm. 
* The minimum value of depth also indicates a possible data input error with the value being a negative number. 


```{r data cleansing, echo=TRUE, warning=FALSE, include=TRUE}
#Output descriptive statistics for all variables
summary(dfraw)

#Remove missing data
dfraw <- na.omit(dfraw)

#Sort data by Diameter
dfraw <- dfraw[order(-dfraw$DIAMETER_MM),]
head(dfraw, 5)

#Remove observations with diameters above maximum set diameter
maximum_diamater <- 10000

cleandf <- subset(dfraw, DIAMETER_MM < maximum_diamater)

#Remove observations with negative values of depth
cleandf <- subset(cleandf, DEPTH_M > 0)

#Convert JOB_TYPE to a factor and add to the dataset
cleandf$JOB_TYPE <- as.factor(cleandf$JOB_TYPE)
#Ensure JOB_TYPE is a factor data type
class(cleandf$JOB_TYPE)

#Convert MATERIAL to a factor and add to the dataset
cleandf$MATERIAL <- as.factor(cleandf$MATERIAL)
#Ensure MATERIAL is a factor data type
class(cleandf$MATERIAL)
```
# 1. Describe the data provided.

## Visualising the data 
Visualising a dataset with histograms and bar charts often helps get a better understanding of the data. In this case, it appears that the 4 continuous variables all have a positive skew. Diameter appears to have clusters of data which suggests the piping company have preset diameters that are produced. 

```{r descriptive visualising, echo=TRUE, warning=FALSE, include=TRUE}
#Check the descriptive statistics after cleaning the data
summary(cleandf)

#Create histograms to visualise the continuous variables
variables <- colnames(cleandf[c(3,4,6,7)])

for(i in variables){
hist(cleandf[,i],
        xlab = i,
        main = c("Histogram of", i),
        breaks = sqrt(nrow(cleandf))
        )
  }

#Create bar charts to visualise the categorical variables
p1 <- ggplot(cleandf, aes(x = JOB_TYPE, fill = JOB_TYPE)) + 
  geom_bar() +
  theme(legend.position = "none") +
  labs(y = "Frequency", x = "Job Type") +
  ggtitle("Bar chart to show Job Type frequency ")
p1

p2 <- ggplot(cleandf, aes(x = MATERIAL, fill = MATERIAL)) + 
  geom_bar() +
  theme(legend.position = "none") +
  labs(y = "Frequency", x = "Material") +
  ggtitle("Bar chart to show Material frequency ")
p2
```

# 2. Show how the cost of jobs is related to other features in the data.

## Visualising continous variables
Diameter has a clear positive linear relationship with job cost whereas depth and length have far weaker relationships with job cost 

```{r visualising continous relationships, echo=TRUE, warning=FALSE, include=TRUE}
#Plot regression line graph for DEPTH_MM and JOB_COST
p3 <- ggplot(cleandf, aes(x = DEPTH_M, y = JOB_COST)) +
    geom_point() +
    geom_smooth(method = 'lm') 
p3
#Plot regression line graph for DIAMTER_MM and JOB_COST
p4 <- ggplot(cleandf, aes(x = DIAMETER_MM, y = JOB_COST)) +
    geom_point() +
    geom_smooth(method = 'lm')
p4
#Plot regression line graph for LENGTH_M and JOB_COST
p5 <- ggplot(cleandf, aes(x = LENGTH_M, y = JOB_COST)) +
    geom_point() +
    geom_smooth(method = 'lm')
p5
```

## Visualising catagorical variables 

```{r visualising catagorical relationships, echo=TRUE, warning=FALSE, include=TRUE}
p6 <- ggplot(cleandf, aes(x = JOB_TYPE, y = JOB_COST)) +
      geom_bar(stat = "identity")
p6

p7 <- ggplot(cleandf, aes(x = MATERIAL, y = JOB_COST)) +
      geom_bar(stat = "identity")
p7
```

# Analysing the data

## Normalise/Standardise the data
Normalising the predictor variables allows for direct comparison of effect size. 

```{r create z variables, echo=TRUE, warning=FALSE, include=TRUE}
#Z transform the continous predictor variables
cleandf$zDEPTH <- (cleandf$DEPTH_M - mean(cleandf$DEPTH_M))/sd(cleandf$DEPTH_M)

cleandf$zDIAMETER <- (cleandf$DIAMETER_MM - mean(cleandf$DIAMETER_MM))/sd(cleandf$DIAMETER_MM)

cleandf$zLENGTH <- (cleandf$LENGTH_M - mean(cleandf$LENGTH_M))/sd(cleandf$LENGTH_M)
#Check variables have been normalised 
summary(cleandf)

#Write cleandf
write.table(cleandf, file = paste("cleanData.csv", sep = "/"), sep = ",", row.names = FALSE)
```

## Examine relationships 

```{r analysing relationships, echo=TRUE, warning=FALSE, include=TRUE}
#Multiple regression analysis
JOB_COSTlm <- lm(JOB_COST ~ zDEPTH + zDIAMETER + zLENGTH + JOB_TYPE, data=cleandf)
#Examine regression analysis
summary(JOB_COSTlm)

#Multiple regression analysis
JOB_COSTlm2 <- lm(JOB_COST ~ zDEPTH + zDIAMETER + zLENGTH + JOB_TYPE + MATERIAL, data=cleandf)
#Examine regression analysis
summary(JOB_COSTlm2)

#Compare models
anova(JOB_COSTlm, JOB_COSTlm2)
```
 
## Examine interactions 

```{r analysing interactions, echo=TRUE, warning=FALSE, include=TRUE}
#Regression analysis including interactions
JOB_COSTlm3 <- lm(JOB_COST ~ zDEPTH + zDIAMETER + zLENGTH + JOB_TYPE + MATERIAL + zDEPTH:JOB_TYPE + zDIAMETER:JOB_TYPE + zLENGTH:JOB_TYPE, data=cleandf)
#Examine regression analysis
summary(JOB_COSTlm3)
```

# 3. Prediciting the cost of a job 
regression analysis without normalised variables 

```{r predicting job cost , echo=TRUE, warning=FALSE, include=TRUE}
#Regression analysis predicting job cost 
JOB_COSTlm4 <- lm(JOB_COST ~ DEPTH_M + DIAMETER_MM + LENGTH_M + JOB_TYPE + MATERIAL, data=cleandf)
#Examine regression analysis
summary(JOB_COSTlm4)
```

The model explains **81.4%** of variance in the cost of a job. This suggest there are other variables effecting job cost. 81.4% is still a considerable amount of variance explained and so developing an equation to predict job cost may be useful. 

### Regression formula

Using the formula for a regression analysis we can create an equation to predict the cost of a job 
**y = b1 * X1 + b2 * x2 + ... + bk * xk + a**

### Formula for the current model

Job Cost = DEPTH_M * DEPTH_M estimate + DIAMETER_MM * DIAMETER_MM estimate + LENGTH_M * LENGTH_M estimate + (1 if JOB_TPYE = replace, if not 0) * JOB_TYPEReplace estimate + (1 if MATERIAL = CI, if not 0) * MATERIALCI estimate + (1 if MATERIAL = ST, if not 0) * MATERIALST estiamte + intercept

Job Cost = DEPTH_M * 3876.00 + DIAMETER_MM * 75.89 + LENGTH_M * 182.20 + (1 if JOB_TPYE = replace, if not 0) * 8170.00 + (1 if MATERIAL = CI, if not 0) * -40.52 + (1 if MATERIAL = CI, if not 0) * -34.74 - 20550.00

### Formula in action 

We can use this equation to calculate the job cost of a typical job if the piping company started creating piping in a new larger diameter size of 600mm.

Using values of (mean values used)
* DEPTH_M = 2.673
* DIMATER_MM = 600
* LENGTH = 25.13 
* JOB_TYPE = Replace
* MATERIAL = AC 


```{r predicting job cost with large diameter , echo=TRUE, warning=FALSE, include=TRUE}
JobCost = (2.673 * 3876) + (600 * 75.89) + (25.13 * 182.2) + (1 * 8170) + (0 * -40.52) + (0 * -34.74) - 20550

JobCost
```