---
title: "Pipes"
author: "Alex Montgomery"
date: "10/03/2022"
output: html_document
---

```{r libraries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #Load tidyverse
library(ggplot2) #Load ggplot2
```

```{r overwatch logo, out.width = "50%", fig.align = "center", echo=FALSE}
#Load and display pipe fitting jpeg
img1 <- "images/professional-pipe-fitter.jpg"
knitr::include_graphics(img1)
```

## Pipes Dataset
The following dataset that shall be explored contains work orders from a utility company. The data represents jobs performed on pipes belonging to the company.

```{r load data, echo=TRUE, warning=FALSE, include=TRUE}
#Load the data set
df = read.csv("data/raw/example.work.orders.csv")
#Check the data is as expected
head(df, 5)
```

## Codebook
Variable Title | Explanation |
|:------:|:------|
JOB_ID | The unique identity of the work order |
JOB_TYPE | The type of job performed |
DEPTH_M | The depth of the pipe (m) |
DIAMETER MM | The diameter of the pipe (mm) |
MATERIAL | The material of the pipe |
LENGTH_M | The length of the pipe (m) |
JOB_COST | The cost of the job (£) |

## Descriptive Statistics


